---
title: "A metaframe codebook"
author: "Carl Frederick"
date: "August 31, 2015"
output: html_document
---

These are my notes to self about interacting with data.frame attributes in R. 
The ultimate goal is to develop a working model for the meta-frame package and
writing a codebook function that access the codebook information.

**Note: this file won't render completely, but the codebook will**

# Setup

I will create a fake data set to practice on:

```{r fake data }
  fake <- data.frame(x = floor(runif(100, 0, 101)))
  fake$y = fake$x * .35 + rnorm(100, 0,5)
  fake$c = letters[((fake$x %% 26) + 1)]
  fake$f = LETTERS[((fake$x %% 5) + 1)]
  fake$f <- factor(fake$f)
 
  #add missing data
  set.seed(8501)
  
  fake$y[sample(1:100, 10)] <- NA
  fake$c[sample(1:100, 17)] <- NA
  fake$f[sample(1:100, 13)] <- NA
```

# Metaframe object structure

Based on our discussions, we decided that it might be best if all the metaframe
attributes were saved in a single object.  Lets try out an object with the following
structure:

```{r metaframe structure}
  metaframe <- NULL
  metaframe$source <- list("this is the source info for the data.frame as a whole", 
                           x = "http://fake.url", 
                           y = "calculated",   
                           c = "I made this up")
  metaframe$units <- list(x = "units sold",
                          y = "1925 Dollars")
  metaframe$description <- list("Fictious data to test accessing attributes and writing a codebook function",
                                x = "Number of units sold last year",
                                y = "Net profits",
                                c = "Salesperson initial",
                                f = "Model Type")
 
  attr(fake, "metaframe") <- metaframe
  fake <- structure(fake, class = c("metaframe", "data.frame"))
```


# Simple codebook function

In the end, I want the following structure for a codebook:

----

> **Column:** colname  
> **Label:**  metaframe\$description\$colname  
> **Type:**   class  
> **Source:** metaframe\$source\$colname  
> **Units:**  metaframe\$units\$colname  

**Description**  
`metaframe$annotation$colname`  
`metaframe$revision$colname`

**Summary**  
if numeric then mean, sd, quantiles, num.missing
if char/factor then table of (5/10? most common values), 

----


```{r codebook}
codebook <- function(mf.obj, file=NULL, replace=FALSE, render=FALSE) {
  if (!inherits(mf.obj, "metaframe")) {
    stop("not a metaframe object")
  }
  varnames <- names(mf.obj)
  nvars <- length(varnames)
  meta_data <- attr(mf.obj, "metaframe")
  mf.name <- deparse(substitute(mf.obj))
  
  if (is.null(file)) {
    fname <- tempfile("codebook", fileext = ".Rmd")
  } else {
    if (file.exists(file)) {
      if (replace) {
        message(paste("Overwriting", file))
        file.create(file)
      } else {
        stop(paste("File", file, "already exists. Set replace = TRUE to overwrite"))
      }
    } else {
      file.create(file)
    }
    fname <- file(description = file, open = "r+")
  }
  
  header <- c("---", paste('title: "Codebook for `', mf.name, '`"', sep=""), paste("date:", date()),
      "output: ", "  html_document:", "    number_sections: no", "    toc: yes", "    toc_depth: 2", "---", "\n")
  
  cat(header, sep="\n", file=fname, append=FALSE)

  
  #Check for Data-level info
  if (any(unlist(sapply(meta_data, names))=="")) {
    varnames <- c("", varnames)
    nvars <- nvars + 1
  }
  
  #Cycle through (remember the two trailing spaces!)
  for (i in 1:nvars) {
    idx <-  names(meta_data$description)==varnames[i]
    cb.name <-   paste("## `", varnames[i], "`  \n\n")
    cb.label <-  paste("> **Label:  **", meta_data$description[idx], "  ")
    cb.type <-   paste("> **Type:   **", ifelse(varnames[i]=="",  class(mf.obj), class(mf.obj[,varnames[i]])), "  ")
    cb.source <- paste("> **Source: **", meta_data$source[idx], "  ")
    cb.unit <-   paste("> **Units:  **", meta_data$units[idx], "  ")
    cb.desc <-   paste(ifelse(varnames[i]=="", 
                              paste("**Metaframe Description**  \n", meta_data$description[idx], "  \n\n"), 
                              "**Description**  \n"),
                       ifelse(is.null(meta_data$annotation[idx]), 
                              "No annotations.  \n\n", 
                              paste(meta_data$annotation[idx], "\n\n")),
                       ifelse(is.null(meta_data$revision[idx]), 
                              "No revisions.  \n\n", 
                              paste(meta_data$revision[idx], "\n\n")),
                       ifelse(varnames[i]=="", paste("Source: ", meta_data$source[idx], "  \n"), ""))
    
    if (varnames[i]=="") {
      cat(c(cb.desc, "\n"), 
          sep="\n", file=fname, append=TRUE)
    } else  {
      cat(c("---- \n", 
            cb.name, cb.label, cb.type, cb.source, cb.unit, "\n", 
            cb.desc, "\n",
            "**Summary**  \nInfo in here", "\n"), sep="\n", file=fname, append=TRUE)
    }
  }

  if (is.null(file)) {
    if (render) browseURL(rmarkdown::render(fname))
    unlink(fname)
  }
  if (!is.null(file)) {
    if (render) browseURL(rmarkdown::render(file))
    close(fname)
  }
}

#debug(codebook)
codebook(fake, render=TRUE)

```

The codebook should have opened in your browser!


